---
title: "TRANSLATOR Co-occurrence baseline calculations"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
DEBUG = FALSE
```

## Introduction

This document gives descriptive statistics on our pilot relation annotation project.
It also establishes baseline performance of co-occurrence for the six relations.
These will be used as benchmarks in development of relation extraction applications during the next funding period.

Author: Kevin Cohen

```{r function.definitions, echo=FALSE}
# Countables for *every* relation
yes = 0;
no = 0;
anaphora = 0;
entity = 0;
other.relations = 0;
segmentation = 0;
speculation = 0;

counts.relation <- function(relation, yeses, nos, anaphora, entities, other.relations, segmentation) {
  this.relation <- as.data.frame(relation = relation, 
                                 yes = yeses,
                                 no = nos,
                                 anaphora = anaphora,
                                 entities = entities,
                                 other.relations = other.relations,
                                 segmentation = segmentation)
  return(this.relation)
}


#a.relation <- counts.relation("gene-gene", yes, no, anaphora, entity, other.relations, segmentation)

# need a separate function for the directional ones
calculateTruePositives <- function(results.df) {
  true_positives <- sum(results.df$yes, results.df$increase, results.df$decrease, results.df$direction.unspecified)
  return(true_positives)
}
  

calculateFalseNegatives <- function(anaphora) {
  return(anaphora)
}

calculatePrecision <- function(true_positive, false_positive) {
  if(DEBUG) { print(paste("TPs:", true_positive, "FPs:", false_positive)) }
  precision = true_positive / (true_positive + false_positive)
  if (DEBUG) { print(paste("Precisions:", precision)) }
  return(precision)
}
calculateRecall <- function(true_positive, false_negative) {
  recall = true_positive / (true_positive + false_negative)
  return(recall)
}
calculateF_measure <- function(precision, recall) {
  f_measure = (2 * precision * recall) / (precision + recall)
  return(f_measure)
}

# TESTS

```

## Calculate figures of merit for all relations



```{r read.in.data, echo=FALSE}

# development data
# relations = c("chemical.gene", "gene.gene", "disease.symptom")
# yeses = c(100, 25, 100)
# nos = c(50, 175, 25)
# anaphora = c(10, 20, 30)
# entity = c(25, 50, 10)
# speculations = c(10, 15, 5)
# other_relations = c(10, 10, 10)
# precisions = c(0, 0, 0)
# recalls = c(0, 0, 0)
# f_measures = c(0, 0, 0)

# if trouble, see the code for sentence-level binary classification
baseline_data <- read.table("/Users/kevincohen/Dropbox/Scripts-new/TRANSLATOR baselines - Sheet1.tsv",
                            header = TRUE, sep = "\t")
# for initial dev data
#baseline_data <- data.frame(relation = relations, yes = yeses, no = nos, anaphora = anaphora, entity = entity, speculation = speculations, 
#                            other.relations = other_relations, precision = precisions, recall = recalls, f_measure = f_measures)
#head(baseline_data)
print.data.frame(baseline_data)
#true_positives = baseline_data$yes
true_positives = calculateTruePositives(baseline_data)

#true_positives
#false_positives <- baseline_data$no + baseline_data$other_relations # NO IDEA why this isn't working!!!
false_positives <- sum(baseline_data$no, baseline_data$other_relations)


if (DEBUG) {
    baseline_data$no
    baseline_data$other.relations
    false_positives
}
false_negatives = calculateFalseNegatives(baseline_data$anaphora)
#false_negatives
baseline_data$precision = calculatePrecision(true_positives, false_positives)
baseline_data$recall = calculateRecall(true_positives, false_negatives)
baseline_data$f_measure = calculateF_measure(baseline_data$precision, baseline_data$recall)
print.data.frame(baseline_data)
#head(baseline_data)
#print.data.frame(cbind(baseline_data[,1], baseline_data[,9:11]))
#performance_to_print.df <- cbind(baseline_data[1,], baseline_data[,9:11])

#relations <- baseline_data[, baseline_data$relation]
#precisions <- baseline_data[,baseline_data$precision]
#recalls <- baseline_data[,baseline_data$recall]
#f_measures <- baseline_data[baseline_data$f_measure]
#performance_to_print.df <- cbind(precisions, recalls, )
#print.data.frame(performance_to_print.df)
```

## Descriptive statistics of dataset

_TODO: plot balance per relation_
```{r, echo=FALSE}
total_exemplars = sum(baseline_data[, c(2:ncol(baseline_data))])
#total_exemplars
baseline_data$total_exemplars <- sum(baseline_data[2:ncol(baseline_data)]) # not what I want
#baseline_data$total_exemplars
#my_plot_statistics <- ggplot(data = baseline_data, aes(x = relation, y = sum(c(yes, no, anaphora, entity, other.relations)) + geom_bar(stat="identity")))
#my_plot_statistics <- ggplot(data = baseline_data, aes(x = relation, y = no)) + geom_bar(stat="identity")
#my_plot_statistics <- ggplot(data = baseline_data, aes(x = relation, y = yes + no + anaphora + entity + other.relations)) + geom_bar(stat="identity")
my_plot_statistics <- ggplot(data = baseline_data, aes(x = relation, y = yes + no + anaphora + entity + other.relations)) + geom_bar(stat="identity") + xlab("Relation") + ggtitle("Total annotations minus sentence segmentation errors") + ylab("Exemplars")
my_plot_statistics 
```

## Plot values for figures of merit

```{r, echo=FALSE}
#library(ggplot2)
library(cowplot)
# change the name of the argument--can be used for *any* dataframe that has the appropriate columns...
plot_figures_of_merit <- function(baseline_data) {
  my_plot_f <- ggplot(data = baseline_data, aes(x = relation, y = f_measure)) + geom_boxplot() + ylab("F-measure") + ylim(0, 1.0) + ggtitle("Baseline by relation: F-measure")
 # my_plot_f

  my_plot_p <- ggplot(data = baseline_data, aes(x = relation, y = precision)) + geom_boxplot() + ylab("Precision") + ylim(0, 1.0) + ggtitle("Baseline by relation: Precision")
#  my_plot_p

  #my_plot_r <- ggplot(data = baseline_data, aes(y = f_measure)) + geom_boxplot() + ylab("F-measure") + ylim(0, 1.0) + ggtitle("Variability across relations: F-measure")
  #my_plot_r

  my_plot_r <- ggplot(data = baseline_data, aes(x = relation, y = recall)) + geom_boxplot() + ylab("Recall") + ylim(0, 1.0) + ggtitle("Baseline by relation: Recall")
 # my_plot_r

plot_grid(
  my_plot_p,
  my_plot_r,
  my_plot_f,
  labels = "AUTO", 
  ncol = 1
  #nrow = nrow(baseline_data)
  #ncol = 2, 
  #nrow = 2
  #ncol = 2, nrow = 2
)
}

plot_figures_of_merit(baseline_data)

precision_all_relations <- ggplot(data = baseline_data, aes(y = precision)) + geom_boxplot() + ylab("Precision") + ylim(0, 1.0) + ggtitle("Variability across relations: Precision")
precision_all_relations
recall_all_relations  <- ggplot(data = baseline_data, aes(y = recall)) + geom_boxplot() + ylab("Recall") + ylim(0, 1.0) + ggtitle("Variability across relations: Recall")
#recall_all_relations
f_measure_all_relations  <- ggplot(data = baseline_data, aes(y = f_measure)) + geom_boxplot() + ylab("F-measure") + ylim(0, 1.0) + ggtitle("Variability across relations: F-measure")
#f_measure_all_relations

plot_grid(
  precision_all_relations,
  recall_all_relations,
  f_measure_all_relations,
  labels = "AUTO",
  ncol = 2
)
```


```{r}

```


```{r}

```

